---
title: "Schedule"
description: |
  Schedule for Math 58B, Introduction to Biostatistics, Spring 2023
execute: 
  message: false
  warning: false
---

```{r}
#| echo: false
library(tidyverse)
library(kableExtra)
library(knitr)
```

Here's your roadmap for the semester!
Each week, follow the general process outlined below:

- Before class on Tuesday, read the suggested article {{< fa newspaper >}} <i class="fas fa-newspaper"></i>

- Enjoy the notes / readings <i class="fas fa-book-open"></i>  

- Attend class, review the warm-up <i class="fas fa-pencil-alt"></i> if you have any questions after completing it during class.

- Complete the Lab <i class="fas fa-desktop"></i> and HW <i class="fas fa-laptop"></i> assignments 
   - Lab due every Tuesday by 11:59pm
   - HW due every Thursday by 11:59pm

- Discuss the reflection questions <i class="fas fa-lightbulb" target="_blank"></i> and ethics considerations <i class="fas fa-balance-scale"></i> (see class notes) with your classmates, mentor, and professor  

- IMS: <a href = "https://openintro-ims.netlify.app/" target= "_blank"><b>Introduction to Modern Statistics</b></a> by Çetinkaya-Rundel and Hardin.

- ISCAM: <a href = "http://www.rossmanchance.com/iscam3/" target = "_blank"><b>Investigating Statistics, Concepts, Applications, and Methods</b></a> by Chance and Rossman.  Many of the examples come from ISCAM, a pdf of which can be purchased for $5.
   - <a href = "http://www.rossmanchance.com/applets/index2021.html" target = "_blank">ISCAM applets</a>
   - <a href = "https://www.lock5stat.com/StatKey/" target = "_blank">StatKey (boostrapping) applets</a>

```{r}
#| echo: false
schedule <- tribble(
  ~`date`, ~topic, ~agenda, ~readings, ~`article (Tues)`, ~assignments,
  "Week 1 <br> 1.17.23", 'Intro + <br> variables &  <br> studies <br> SLR',"• course info <br> • tests <br> • studies <br> • causation", 
  '<a href="https://st47s.com/Math58/Notes/intro.html#intro" target = "_blank"><i class="fas fa-book-open"></i> Introduction</a> <br> <a href="https://st47s.com/Math58/Notes/intro.html#experim" target = "_blank"><i class="fas fa-book-open"></i> Studies</a>  <br> <a href = "http://www.rossmanchance.com/iscam3/" target = "_blank"><i class="fas fa-book-open"></i> ISCAM 1.1</a> <br> <a href="https://openintro-ims.netlify.app/data-hello.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 1 + 2</a>',
  '<i class="fas fa-newspaper"></i> none <br> <br> <a href = "https://hardin47.github.io/CURV/scholars/enders.html" target = "_blank">Felicity Enders</a>',
  '<a href="https://m58-intro-stats.netlify.app/slides/wu1_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU1</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/wu2_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU2</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW0s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW0 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW0s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW0 Rmd</a>',
  
  
  "Week 2 <br> 1.24.23", 'Correlation + <br>Least<br>squares<br> ',"• correlation<br> • linear model <br> • R^2", 
  '<a href="https://st47s.com/Math58/Notes/regdesc.html#cor" target = "_blank"><i class="fas fa-book-open"></i> Cor</a> <br> <a href="https://st47s.com/Math58/Notes/regdesc.html#ls" target = "_blank"><i class="fas fa-book-open"></i> Least Sq</a>  <br> <a href = "http://www.rossmanchance.com/iscam3/" target = "_blank"><i class="fas fa-book-open"></i> ISCAM 5.6, 5.7, 5.8</a> <br> <a href="https://openintro-ims.netlify.app/model-slr.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 7</a> <br> <a href = "http://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html" target = "_blank"><i class = "fas fa-gamepad"></i> guess corr</a> <br> <a href = "http://www.rossmanchance.com/applets/2021/regshuffle/regshuffle.htm" target = "_blank"><i class = "fas fa-gamepad"></i> least sq</a>',
  '<a href="https://www.wired.com/story/the-astrazeneca-covid-vaccine-data-isnt-up-to-snuff/" target = "_blank"><i class="fas fa-newspaper"></i> AstraZeneca</a> <br> <br> <a href = "https://hardin47.github.io/CURV/scholars/dubois.html" target = "_blank">W.E.B. Du Bois</a>',
  '<a href="https://m58-intro-stats.netlify.app/slides/wu3_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU3</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/wu4_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU4</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW1s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW1 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW1s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW1 Rmd</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab1_m58_s23_intro_to_data.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab1 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab1_m58_s23_intro_to_data.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab1 Rmd</a>',
  
  "Week 3 <br> 1.31.23", 'Hyp Test + <br>Randomization<br>Test<br> ',"•  structure of<br> hypothesis<br> testing", 
  '<a href="https://st47s.com/Math58/Notes/foundations-for-inference.html#randtest" target = "_blank"><i class="fas fa-book-open"></i> Tests</a>  <br> <a href = "http://www.rossmanchance.com/applets/2021/chisqshuffle/ChiSqShuffle.htm?twobytwo=1" target = "_blank"><i class = "fas fa-gamepad"></i> 2x2 Rand</a> <br> <a href="https://openintro-ims.netlify.app/foundations-randomization.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 11</a>',
  '<a href="https://gizmodo.com/i-fooled-millions-into-thinking-chocolate-helps-weight-1707251800" target = "_blank"><i class="fas fa-newspaper"></i> chocolate</a> <br> <br> <a href = "https://hardin47.github.io/CURV/scholars/nuzzo.html" target = "_blank">Regina Nuzzo</a>',
  '<a href="https://m58-intro-stats.netlify.app/slides/wu5_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU5</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/wu6_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU6</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW2s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW2 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW2s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW2 Rmd</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab2_m58_s23_2quant.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab2 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab2_m58_s23_2quant.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab2 Rmd</a>',
  
  "Week 4 <br> 2.7.23", 'Bootstrapping',"• boot samp dist <br> • boot CI ", 
  '<a href="https://st47s.com/Math58/Notes/foundations-for-inference.html#boot" target = "_blank"><i class="fas fa-book-open"></i> Bootstrapping</a> <br> <a href="https://st47s.com/Math58/Notes/foundations-for-inference.html#bootCI" target = "_blank"><i class="fas fa-book-open"></i> Boot CIs</a>  <br> <a href = "https://www.lock5stat.com/StatKey/" target = "_blank"><i class="fas fa-gamepad"></i> StatKey </a> <br> <a href = "http://www.rossmanchance.com/applets/2021/confsim/ConfSim.html" target = "_blank"><i class = "fas fa-gamepad"></i> CIs</a> <br> <a href="https://openintro-ims.netlify.app/foundations-bootstrapping.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 12</a>',
  '<a href="https://www.nytimes.com/2021/01/27/well/move/the-best-time-of-day-to-exercise.html" target = "_blank"><i class="fas fa-newspaper"></i> confounding</a> <br> <br> <a href = "https://hardin47.github.io/CURV/scholars/blackwell.html" target = "_blank">David Blackwell</a>',
    '<a href="https://m58-intro-stats.netlify.app/slides/wu7_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU7</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/wu8_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU8</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW3s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW3 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW3s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW3 Rmd</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab3_m58_s23_infer.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab3 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab3_m58_s23_infer.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab3 Rmd</a>',
  
  "Week 5 <br> 2.14.23", 'normality +<br>CLT',"• normal dist<br> • CLT<br> • Z-score <br> • empirical rule <br> • conf int <br> • norm prob <br> • hyp test", 
  '<a href="https://st47s.com/Math58/Notes/foundations-for-inference.html#CLT" target = "_blank"><i class="fas fa-book-open"></i>CLT </a> <br> <a href="https://st47s.com/Math58/Notes/foundations-for-inference.html#norm" target = "_blank"><i class="fas fa-book-open"></i> Norm dist</a>  <br> <a href = "http://www.rossmanchance.com/applets/2021/sampling/OneSample.html" target = "_blank"><i class="fas fa-gamepad"></i> Samp Dist </a> <br> <a href="https://openintro-ims.netlify.app/.html" target = "_blank"><i class="fas fa-book-open"></i> IMS </a>',
  '<a href="https://www.nytimes.com/2020/12/13/learning/what-does-95-effective-mean-teaching-the-math-of-vaccine-efficacy.html" target = "_blank"><i class="fas fa-newspaper"></i> efficacy</a> <br> <a href="https://www.yalemedicine.org/news/covid-19-vaccine-comparison" target = "_blank"><i class="fas fa-newspaper"></i> vaccines</a> <br> <br> <a href = "https://hardin47.github.io/CURV/scholars/nightingale.html" target = "_blank">Florence Nightingale</a>',
      '<a href="https://m58-intro-stats.netlify.app/slides/wu9_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU9</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/wu10_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU10</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW4s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW4 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW4s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW4 Rmd</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab4_m58_s23_BS.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab4 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab4_m58_s23_BS.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab4 Rmd</a>',
  
  "Week 6 <br> 2.21.23", 'Errors + <br>Power +<br>Sampling',"• Type I error <br> • Type II error <br> • Power <br> • biased samples", 
  '<a href="https://st47s.com/Math58/Notes/foundations-for-inference.html#errors" target = "_blank"><i class="fas fa-book-open"></i> Errors</a> <br> <a href="https://st47s.com/Math58/Notes/foundations-for-inference.html#samp" target = "_blank"><i class="fas fa-book-open"></i> Sampling</a>  <br> <a href = "http://www.rossmanchance.com/applets/2021/power/power.html" target = "_blank"><i class="fas fa-gamepad"></i> Power</a> <br> <a href="https://openintro-ims.netlify.app/decerr.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 14 + 2</a>',
  '<a href="https://www.nytimes.com/2022/01/01/upshot/pregnancy-birth-genetic-testing.html" target = "_blank"><i class="fas fa-newspaper"></i> type II</a> <br> <a href="https://www.forbes.com/sites/ellenmatloff/2022/01/06/what-the-nytimes-got-wrong-on-prenatal-screening/?sh=1def5aa137a7" target = "_blank"><i class="fas fa-newspaper"></i> type I</a> <br> <br> <a href = "https://hardin47.github.io/CURV/scholars/wright.html" target = "_blank">Tommy Wright</a>',
        '<a href="https://m58-intro-stats.netlify.app/slides/wu11_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU11</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/wu12_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU12</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW5s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW5 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW5s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW5 Rmd</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab5_m58_s23_norm.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab5 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab5_m58_s23_norm.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab5 Rmd</a>',
  
  "Week 7 <br> 2.28.23", 'Rel Risk +<br>Odds Ratios',"•  RR <br> • OR<br> • Case-control <br> • Cohort ", 
  '<a href="https://st47s.com/Math58/Notes/inference-for-categorical-data.html#diffprop" target = "_blank"><i class="fas fa-book-open"></i> diff props</a> <br> <a href="https://st47s.com/Math58/Notes/inference-for-categorical-data.html#rr" target = "_blank"><i class="fas fa-book-open"></i> RR</a>  <br> <a href="https://st47s.com/Math58/Notes/inference-for-categorical-data.html#or" target = "_blank"><i class="fas fa-book-open"></i> OR</a>  <br> <a href = "http://www.rossmanchance.com/iscam3/" target = "_blank"><i class="fas fa-book-open"></i> ISCAM 3.9-3.11</a> <br> <a href="https://openintro-ims.netlify.app/inference-two-props.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 17</a>',
  '<a href="https://www.covid-datascience.com/post/israeli-data-how-can-efficacy-vs-severe-disease-be-strong-when-60-of-hospitalized-are-vaccinated" target = "_blank"><i class="fas fa-newspaper"></i> interaction</a> <br> <br> <a href = "https://hardin47.github.io/CURV/scholars/jackson.html" target = "_blank">Monica Jackson</a> ',
  '<a href="https://m58-intro-stats.netlify.app/slides/wu13_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU13</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/wu14_ma58b_s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> WU14</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW6s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW6 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW6s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW6 Rmd</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab6_m58_s23_errors.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab6 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab6_m58_s23_errors.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab6 Rmd</a>',

  'Week 8 <br> 3.7.23', 'catch-up',
  "",
  '<i class="fas fa-book-open"></i> study materials <br> (see Canvas) <br> <a href="https://m58-intro-stats.netlify.app/slides/adviceExam1_s23.docx" target = "_blank"><i class="fas fa-pencil-alt"></i>advice for exam 1</a>',
  "no lab on 3.10.23",
  '<a href="https://m58-intro-stats.netlify.app/slides/MA58_HW7s23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW pdf - not due ever</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HW7s23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW Rmd - not due ever</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab7_m58_s23_RROR.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab7 pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/lab7_m58_s23_RROR.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> Lab7 Rmd</a>',
  
  
'<font color = "#114B5F"><b>3.9.23</b></font>', '<font color = "#114B5F"><b>Exam 1</b></font>',"", 
  '',
  '',
  '',    

  '3.14.23', 'spring break',"", 
  '',
  '',
  '', 
  
  
  "Week 9 <br> 3.21.23", 'χ^2 Test',"• expected counts <br> • X^2 statistic <br> • χ^2 test", 
  '<a href="https://st47s.com/Math58/Notes/inference-for-categorical-data.html#chisq2" target = "_blank"><i class="fas fa-book-open"></i> χ^2 Test</a>  <br> <a href = "http://www.rossmanchance.com/iscam3/" target = "_blank"><i class="fas fa-book-open"></i> ISCAM 3.2</a> <br> <a href="https://openintro-ims.netlify.app/inference-tables.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 18</a>',
  '<a href="https://wwwnc.cdc.gov/eid/article/26/11/20-3003_article" target = "_blank"><i class="fas fa-newspaper"></i> masks</a> <br> ',
  '',
  
  "Week 10 <br> 3.28.23", '1 mean +<br>pred interval',"• dist of X-bar <br> • SE(X-bar) <br> • t-test of μ<br> • t-CI for μ <br> • PI for X_{n+1}", 
  '<a href="https://st47s.com/Math58/Notes/inference-for-numerical-data.html#mean1inf" target = "_blank"><i class="fas fa-book-open"></i> 1 mean</a> <br> <a href="https://st47s.com/Math58/Notes/inference-for-numerical-data.html#mean1inf" target = "_blank"><i class="fas fa-book-open"></i> pred int</a>  <br> <a href="https://openintro-ims.netlify.app/inference-one-mean.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 19</a>',
  '<a href="https://static1.squarespace.com/static/5d2633cb0ef5e4000134fa02/t/5d72cb394452727586b69c2e/1567804217938/policy_brief_indigenous_data_sovereignty_in_the_united_states.pdf" target = "_blank"><i class="fas fa-newspaper"><br> </i> Data sovereignty</a> <br> no lab on 3.31.23',
  '<a href="https://m58-intro-stats.netlify.app/slides/MA58_HWs23.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> HW pdf</a> <br> <a href="https://m58-intro-stats.netlify.app/slides/MA58_HWs23.Rmd" target = "_blank"><i class="fas fa-pencil-alt"></i> HW Rmd</a> <br> Pilot Study Due',

'<font color = "#114B5F"><b>3.28.23</b></font>', '<font color = "#114B5F"><b>Proj 1: Pilot</b></font>',"", 
  '','',
  '<a href="https://m58-intro-stats.netlify.app/slides/Pilot_m58_s22_Islands.docx" target = "_blank"><i class="fas fa-desktop"></i> Proj 1: Pilot</a>',
  
  "Week 11 <br> 4.4.23", '2 means +<br>t-proced',"• dist of X-bar_1 - X-bar_2 <br> • SE(X-bar_1 - X-bar_2) <br> • t-test of μ_1 - μ_2 <br> • t-CI for μ_1 - μ_2", 
  '<a href="https://st47s.com/Math58/Notes/inference-for-numerical-data.html#mean2inf" target = "_blank"><i class="fas fa-book-open"></i> 2 means</a>  <br> <a href = "http://www.rossmanchance.com/applets/index2021.html" target = "_blank"><i class="fas fa-gamepad"></i> 2 samples</a> <br> <a href="https://openintro-ims.netlify.app/inference-two-means.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 20</a>',
  '<a href="http://www.nytimes.com/2010/09/19/health/research/19trial.html?hp" target = "_blank"><i class="fas fa-newspaper"></i> clinical trials</a>',
  '',

'<font color = "#114B5F"><b>4.4.23</b></font>', '<font color = "#114B5F"><b>Proj 2: Power Analysis</b></font>',"", 
  '','',
  '<a href="https://m58-intro-stats.netlify.app/slides/Power_m58_s22_Islands.docx" target = "_blank"><i class="fas fa-desktop"></i> Proj 2: Power Analysis</a>',

  "Week 12 <br> 4.11.23", 'ANOVA',"• MSG <br> • MSE <br>• F-test ", 
  '<a href="https://st47s.com/Math58/Notes/inference-for-numerical-data.html#ANOVA" target = "_blank"><i class="fas fa-book-open"></i> ANOVA</a>  <br> <a href="https://st47s.com/Math58/Notes/reginf.html#infbeta1" target = "_blank"><i class="fas fa-book-open"></i> Inf on β_1</a>  <br> <a href = "http://www.rossmanchance.com/applets/2021/regshuffle/regshuffle.htm" target = "_blank"><i class="fas fa-gamepad"></i> slopes</a> <br> <a href="https://openintro-ims.netlify.app/inference-many-means.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 22</a> <br> <a href="https://openintro-ims.netlify.app/inf-model-slr.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 24</a>',
  '<a href="https://www.buzzfeednews.com/article/stephaniemlee/brian-wansink-cornell-p-hacking" target = "_blank"><i class="fas fa-newspaper"></i> p-hacking</a> <br> ',
  '',
  
  "Week 13 <br> 4.18.23", 'Inference β_1 + <br>Tech cond + <br>MLR',"•  SE(β_1) <br> • residual plots <br> • multiple X", 
  '<a href="https://st47s.com/Math58/Notes/reginf.html#lmtechcond" target = "_blank"><i class="fas fa-book-open"></i> Tech Cond</a>  <br> <a href = "https://st47s.com/Math58/Notes/reginf.html#MLR" target = "_blank"><i class="fas fa-gamepad"></i> MLR</a> <br> <a href="https://openintro-ims.netlify.app/inf-model-mlr.html" target = "_blank"><i class="fas fa-book-open"></i> IMS 25</a>',
  '<a href="https://jamanetwork.com/journals/jama/fullarticle/208354" target = "_blank"><i class="fas fa-newspaper"></i> 1918 flu</a> <br> ',
  '',

'<font color = "#114B5F"><b>4.18.23</b></font>', '<font color = "#114B5F"><b>Proj 3: Analysis</b></font>',"", 
  '','',
  '<a href="https://m58-intro-stats.netlify.app/slides/PreliminaryAnalysis_m58_s22_Islands.docx" target = "_blank"><i class="fas fa-desktop"></i> Proj 3: Preliminary Analysis</a>',

  "Week 14 <br> 4.25.23", 'catch-up',"", 
  '<i class="fas fa-book-open"></i> study materials <br> (see Canvas) <br> <a href="https://m58-intro-stats.netlify.app/slides/param_est_58.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> when to use what</a>',
  '',
  '',
  
'<font color = "#114B5F"><b>4.27.23</b></font>', '<font color = "#114B5F"><b>Exam 2</b></font>',"", 
  '',
  '',
  '<a href="https://m58-intro-stats.netlify.app/slides/param_est_58.pdf" target = "_blank"><i class="fas fa-pencil-alt"></i> when to use what</a>', 

'<font color = "#114B5F"><b>5.2.23</b></font>', '<font color = "#114B5F"><b>Proj 4: Full draft</b></font>',"", 
  '','',
  '<a href="https://m58-intro-stats.netlify.app/slides/CompleteDraft_m58_s22_Islands.docx" target = "_blank"><i class="fas fa-desktop"></i> Proj 4: Full draft</a>',

'<font color = "#114B5F"><b>5.4 + 12.23</b></font>', '<font color = "#114B5F"><b>Proj Presentation</b></font>',"", 
  '','',
  '<a href="https://m58-intro-stats.netlify.app/slides/Presentation_m58_s22_Islands.docx" target = "_blank"><i class="fas fa-desktop"></i> Proj Presentation</a>',

'<font color = "#114B5F"><b>5.12.23</b></font>', '<font color = "#114B5F"><b>Proj 5: Final Project</b></font>',"", 
  '','',
  '<a href="https://m58-intro-stats.netlify.app/slides/CompleteDraft_m58_s22_Islands.docx" target = "_blank"><i class="fas fa-desktop"></i> Proj 5: Final Project</a>'
  
    )
```


```{r}
#| echo: false
#| column: screen-inset
# can use column: page for the entire page
# can use column: body-outset to bleed outside the center column of text
schedule %>% 
  mutate_all(linebreak) %>% # required for linebreaks to work
  kbl(escape = FALSE, align = c(rep('l',7))) %>%
  kable_styling(bootstrap_options = c("striped", "hover"),
                full_width = TRUE) %>%
  column_spec(1:6,extra_css = "vertical-align:top;")
```




