---
title: "HW 12 -- Math 58B"
author: "your name here"
date: "not due ever"
output: pdf_document
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, fig.height=3, fig.width=5, 
                      fig.align = "center", cache=TRUE)
library(tidyverse)
library(mosaic)
library(broom)
library(praise)
```


### Assignment Summary (Goals)

* Multiple Regression
   - choosing variables
   - confidence intervals for mean response
   - prediction intervals for individual response
   - $R^2$ and adjusted $R^2$

Note:  this HW isn't ever due, so the solutions are already posted.  Also, from Lab 12 and the class notes, there is a lot of R code.  There is virtually no scaffolding of R code in this assignment as a way for you to test yourself with the R code.  But feel free to go look at the solutions if you get stuck!



#### Q2. RailTrails

Consider the RailTrail data available in the mosaic package.

```{r}
# try this to see the variables: ?RailTrail
data("RailTrail")
```

a. Build a linear model with `cloudcover` used to predict the `volume` of riders.  For the process:  make a scatter plot (include a regression line), run the linear model (find the coefficients), check to see if `cloudcover` is significant, and calculate the adjusted $R^2$.


b. Are the technical conditions met for the model?


c. Find a model which predicts `volume` of ridership using some set of the explanatory variables which seem best to you.  Use coefficient p-values as well as adjusted $R^2$ to make your determination.


d. Using model from e., predict the `volume` of ridership for each of the 90 days (i.e., $\hat{Y}$).  Create a plot of residuals vs predictions (add a horizontal line at 0).  Create another plot of predictions ($\hat{Y}$) vs observed value ($Y$) (add a line with intercept 0 and slope 1: `geom_abline(intercept = 0, slope = 1)`)

e.  Did you notice that the residual plot seems to have increasing variability?  let's try a log transformation on the response variable.  In order to compare it to the observed value, we'll need to transform it back (using `exp()`).  Create the linear model using `log(volume)` as the response variable and re-make the residual and prediction plots.


f. **Not on exam 2, might be considered to enhance the project** Which model has a higher adjusted $R^2$, the model with `volume` as the response variable or the model with `log(volume)` as the response variable?


g. **Not on exam 2, might be considered to enhance the project** Find the CI and PI for the `log(volume)` of riders using the model with `avgtemp` and `cloudcover`.  Unfortunately, the intervals need to be generated separately.



```{r}
praise()
```
